|

### 2021-10-01 03-Chp-Systematic-Review.Rmd
End of day:

Following full-text screening, `r n_reports_included` reports describing `r n_included` unique studies met the criteria for inclusion in the review.[@ancelin2012a; @ancelin2013; @arvanitakis2008; @batty2014; @benn2017b; @bettermann2012a; @beydoun2011; @bruce2017; @burgess2017; @chao2015a; @chen2014; @chiang2007; @chitnis2015; @chou2014a; @chuang2015; @cramer2008; @dodge2011; @forti2010; @gnjidic2016; @gottesman2017; @gustafson2012; @haag2009; @hayden2006; @heartprotectionstudycollaborativegroup2002; @hendrie2015; @hippisley-cox2010a; @jick2000; @kimm2011; @kivipelto2001; @kivipelto2005a; @kuo2015; @li2004; @li2005a; @li2010; @liao2013; @liu2019a; @mainousa.g.2005; @mielke2005a; @mielke2010a; @mielke2011; @mukherjee2013; @muller2007; @noale2013; @notkola1998; @ostergaard2017; @pan2018; @parikh2011; @peters2009; @raffaitin2009; @rantanen2017; @rea2005a; @redelmeier2019; @reitz2004a; @reitz2010; @ridker2008; @ronnemaa2011; @schilling2017a; @smeeth2009a; @so2017; @solomon2007; @solomon2009a; @solomon2010; @sparks2008; @strand2013; @su2017; @svensson2019; @szwast2007; @tan2003a; @tynkkynen2016; @tynkkynen2018; @wang2012; @whitmer2005a; @yamada2009b; @yang2015; @yoshitake1995; @zamrini2004; @zandi2005; @zhu2018a; @zimetbaum1992] Table \@ref(tab:studyCharacteristics-table) presents a summary of the characteristics of each study.

@@

__Statins__

@@

__Lipids__

@@

>>>>

Following full-text screening, `r n_included` unique studies (described across `r n_reports_included` reports) met the criteria for inclusion in the review.[@ancelin2012a; @ancelin2013; @arvanitakis2008; @batty2014; @benn2017b; @bettermann2012a; @beydoun2011; @bruce2017; @burgess2017; @chao2015a; @chen2014; @chiang2007; @chitnis2015; @chou2014a; @chuang2015; @cramer2008; @dodge2011; @forti2010; @gnjidic2016; @gottesman2017; @gustafson2012; @haag2009; @hayden2006; @heartprotectionstudycollaborativegroup2002; @hendrie2015; @hippisley-cox2010a; @jick2000; @kimm2011; @kivipelto2001; @kivipelto2005a; @kuo2015; @li2004; @li2005a; @li2010; @liao2013; @liu2019a; @mainousa.g.2005; @mielke2005a; @mielke2010a; @mielke2011; @mukherjee2013; @muller2007; @noale2013; @notkola1998; @ostergaard2017; @pan2018; @parikh2011; @peters2009; @raffaitin2009; @rantanen2017; @rea2005a; @redelmeier2019; @reitz2004a; @reitz2010; @ridker2008; @ronnemaa2011; @schilling2017a; @smeeth2009a; @so2017; @solomon2007; @solomon2009a; @solomon2010; @sparks2008; @strand2013; @su2017; @svensson2019; @szwast2007; @tan2003a; @tynkkynen2016; @tynkkynen2018; @wang2012; @whitmer2005a; @yamada2009b; @yang2015; @yoshitake1995; @zamrini2004; @zandi2005; @zhu2018a; @zimetbaum1992] <!-- TODO implement citation column in and check in rendered version that there are X references -->

Table \@ref(tab:studyCharacteristics-table) presents a summary of the characteristics of each study.

@@

#### Statins

@@

#### Lipids

@@

#### Statins

### 2021-10-01 05-Chp-IPD-Analysis.Rmd
End of day:

---
output:
  bookdown::pdf_document2:
    template: templates/brief_template.tex
  bookdown::word_document2: default
  bookdown::html_document2: default
documentclass: book
---

@@

@@

As part of a broader investigation into the relationship between lipid levels and dementia risk, I sought to

@@

Is there a good way to visualise input of cohorts into an IPD analysis?

Justify analysis based on findings of systematic review - suggestion of variation of effect by age, specifically

@@

Eligible data sources were defined as those that contained lipid measurements and  ipids reported/available as a continuous measure, Data sources which were cross-sectional, either by design or due to the available data (e.g. a study conducted multiple waves, but onyl data from a single wave could be accessed) were excluded. Similarly, due to the limited time-frame, studies making use of population-level electronic health records, which often require an extensive project proposals in order to gain access to the data, were ineligible due to the time and cost involved in applying. The one exception to this was data from the CPRD, which I already had access to as part of the analysis reported in Chapter \ref(cprd-heading).

@@

__Systematic review__

As part of the IPD analysis, eligible observational prospective cohort studies identified through the systematic review detailed in Chapters \@ref(sys-rev-heading) were approached for data access.

In the first instance, the first author on the main study report was emailed (see Appendix \@ref() for a copy of the text and documentation sent). If this did not elict a response, the last/corresponding author was contacted, on the basis that the first author may have been a more junior member.

examining the effect of lipid levels

__Dementia platform UK__

In addition, in an attempt to supplement the analysis with

In addition, to assess the results for  the Dementia Platform UK common-access procedure was utilised. In short, this approach is intended to make it easier to access data from existing dementia cohorts through a centralised application process.

@@

@@

Using multiple imputation to handle missing data is an alternative to a "complete case" approach,[@pigott2001review] where participants missing any covariate are dropped from the dataset. As a sensitivity analysis, I preformed and compared the results of both methods,[@hughes2019] to investigate the impact of multiple imputation on the results.

@@

### Analysis

_Individual patient data approached__

Two stage versus one-stage approach

A two-stage model was used out of necessity, with the one-stage model being precluded by the different datasets being in protected data silos.

__Statistical methods__

Under a two-stage approach, estimates for the effect of each lipid fraction on available dementia outcomes in each study were first calculated for each data source. A common model, adjusted for identical covariates, were obtained using a common model.

To aid interpretability of the outcome, all analyses were standardised by changes in 1-SD of the exposure variable. Clustering within studies was accounted for, given the evidence that ...[@abo-zaid2013]

Hazard ratios were used to quantify the effect of a 1-SD lower exposure to a lipid fraction on dementia outcomes. A discrete proportional hazards model was employed to account for the interval censoring introduced by design of the longitudinal cohort studies.[@wang2017]

The estimates from each data source were then combined using a random-effects meta-analysis

@@

__Investigating effect of patient-level covariates__

@@

__Replicating published findings__

Where an included data source had previously been analysed and results for the association between lipid levels and dementia reported, these were compared to the results of this analysis. If discrepancies were identified, these were investigate.

__Combining published aggregate data and individual participant data__

Where a data source did not provided their individual participant data, but stratified by a patient characteristic of interest in the published data, the agregate date and individual data were combined.

@@

__Author statements on data availability__

@@

Of the XXX studies to which I applied for data access, only three were eventually included in the analysis. Figure \@ref(fig:cohortFlowchart) details whether the cohorts eventually included in the review were identified by the systematic review or via DPUK portal.

In addition, the reasons for cohorts not being included in the analysis are presented. In summary, the requests for data from cohorts identified by the systematic review was characterised by a very low response rate. For the minority who did respond, common reasons given by corresponding authors for not sharing the data included that they no longer worked in the lab, had access to the data, or that they were currently or intended to perform a similar analysis as the one proposed.

For the streamlined DPUK process, where a dedicated project manager liaises with cohort owners on the applicants behalf, the overall response rate was higher, resulting in access to three cohorts. However, even using this dedicated access approach, a response was obtained for less than half of the approach cohorts

@@

Few cohorts were included in both the DPUK and the systematic review sets of cohorts indicating that the DPUK.

As highlighted in the

@@

The four data sources used in this analysis are summarised in Table \@ref(tab:cohortOverview-table) and are described in detail in the following sections.

Of note, all data sources included in the analysis were based in the United Kingdom. This is likely due to the majority of included datasets coming from the Dementia Platform UK route \@ref(fig:cohortFlowchart), which as implied by the name, has a geographical focus on studies performed in Great Britain United Kingdom.

@@

Electronic health record co

Lothian Birth Cohort - 1936

Taiwan, etc.

@@

Several additional cohorts are present in the DPUK collection but were not approach with a request for data acccess. This was primarily due to the information on the online system indicating that the study was relatively new (and so was yet to collect >1 wave of data, as in the case of the XXXX cohort) or the online data dictionary indicating that the exposure variables of interest were not recorded.

Electronic health record co

LBC - 1936

Taiwan etc,

@@

As highlighted in Figure \@ref(fig:cohortFlowchart), several cohorts from the DPUK responding positively but on inspection of data provided these cohorts were excluded.

Several cohorts were excluded on the basis of a lack of exposure variables, including Cam-CAN (had cardiovascular category, but only contained blood pressure).

The reasons for exclusion of these cohorts for which data were provided is illustrated in Table \@ref(tab:dataExcluded-table).

__BRACE__

Likley had dementia at baseline. However, on further investigation of the cohort, it was discovered that

__Memento__

Excluded as criteria for entry was outpatients from memory clinic (unlikely to be dementia free at baseline)

__Generation Scotland__

Cross-sectional data only

__NICOLA__

NICOLA study - only had cross-sectional data

__ELSA__

Data provided but only had ever high cholesterol as a binary variable. Not compatible with

@@

<!----------------------------------------------------------------------------->
(ref:dataExcluded-caption) dataExcluded

(ref:dataExcluded-scaption) dataExcluded

@@

@@

A particularly frustrating in the case of cohorts which had a dedicated data access panel, for example, the Three City Study. Despite multiple attempts to contact the team, there was no response received.

These included studies with  a dedicated access process or data access committee. A cohort of particular interest, having been used in multiple studies identified by the systematic review is the Three City Study in France. This study was approach at multiple points over the curse of this analysis, and no answer was received.

@@

@@

### Data cleaning and harmonisation

Across all cohorts, data cleaning was performed in a similar manner, using commonly named variables, so that a single model could be applied using functional programming.

The one exception to this is the CPRD data, which was held in a different system to the rest.

The advantage of this approach is that it reduces the likelihood of errors in model mis-specification if needing to change variables names from cohort to cohort.

For all cohorts, the first lipid measurement was used for the exposure method,

For heterogeneity across the cohorts, the total time-at-risk was investigated.

@@

# Results

@@

@@

Useful citation for discussion[@levis2021]

@@

<!-- TODO Plot showing when datasets requested were published and comparing this with the -->

This review had a low response rate to requests for data access.

While this is not unexpected, given that a review of IPD studies published between 1987 and 2015 found that fewer than half managed to obtained data from greater than 80% of studies, and that in many cases, the exact percentage of studies for which data were obtained was not accurately reported.
test
There are likely several reasons for this.

* Individual participant data meta-analysis including studies other than randomised controlled trials have less success in obtaining individual participant data from studies.[@nevitt2017a]

* While there is no

For many other reasons also - if

A potential further reason was highlighted during my attendance at the

Letters sent to all cohorts identified through the systematic review 
<!-- TODO will need quite a bit here on the numbers approached and how they responded if at all. Will also need to comment on the additional tactics used.  -->

This is likely due to my junior position as a early career research combined with the impact

Range of reasons why data is not readily available. Privacy concerns, concerns around scoping or "parasitic" behaviour, and a lack of trust (i.e. primary researchers do not trust secondary researchers).<!-- TODO CITATION NEEDED: paper on trust as a big issue + paper on systematic review of barriers to data sharing -->

Unfortunately,

The low response rate and the

Whether or not to press ahead with an IPD analysis in the absence of all (or even most) data is a personal decision, and some have highlighted where they decided not to pursue an IPD study.[@jaspers2014]

@@

For the purposes of this thesis, conducting the IPD was useful as it afforded me the opportunity to experience the methodology.

On a personal level, given the

In addition, it allowed me to analysed two addition previously unanalysed cohort studies which were then utilised in the triangulation study detailed in Chapter \@ref(tri-heading).

@@

#### Opportunistic inclusion of CPRD data

@@

While this analysis did not manage to systematically obtain and analyse a large proporiton of identified data, it did enable an analysis of two previously unanalysed datasets - the CaPS study and the - providing additional data points in

it also attempted to make use

@@

Couple of sections

* Problems with getting data from studies
* Problems with current methods of tracking people down via . Particularly when studies were older, several cases of the address/contact details for the corresponding author being out of date. Matches with decrease in availability as the studies get older.

__Many authors claim that their data can be made “available on request”, despite previous work establishing that these statements are demonstrably untrue in the majority of cases—that when data is requested, it is not actually made available. [@krawczyk2012; @miyakawa2020; @naudet2018] Additionally, previous work found that the availability of data “available on request” declines with article age, indicating that this approach is not a valid long term option for data sharing.[@vines2014]__ <!-- TODO taken from DAS paper, so will need to change language -->

Qualitatively, investigation of a subset of cohorts on an _ad-hoc_ basis reveal that several corresponding/first authors were no longer at the same institution as when the study was reported, and as a result, were unlikely to have access to the institutional email address listed on the study publication. All of these are arguments towards better reporting of data availability, ownership and

Even with the help of the streamlined application process afforded by the DPUK, accessing sufficient data was a challenge. Sharing data has several legal, ethical and logistical challenges, and can often taken several years to obtain. This is in theory what the DPUK was built to address. However, despite claims to a streamlined process, the response rate among cohorts a year after application was 50%. In addition, some cohorts responded that that the proposed study question was already under investigation by another group, and that they would not share the data on this basis. In light of this, a centralised database of ongoing analysis being performed in DPUK would be of enormous help. Finally, the DPUK should more clearly distinguish between those cohorts that are "DPUK native" versus externally hosted, as the response time for externally hosted cohorts is likely to be much longer.

### Study within a review

In future, it could have been worth running a "study within a review", such as that used to examine the best method [@godolphin2019a] to identify and access study data.

Options could include

Previous SWARs have been registered as protocols, but have not yet reported!

@@

#### Aims/Resources

May have been overly ambitious

While it was disappointing that such a small proportion of the available data were obtained for analysis, in terms of project completion, it is probably a good thing.

Given the resource intensity of data cleaning and harmonisation,

@@

>>>>

---
bibliography: bibliography/references.bib
csl: bibliography/nature.csl
output:
  bookdown::word_document2: 
      toc: false
      toc_depth: 3
      reference_docx: templates/word-styles-reference-01.docx
      number_sections: false 
  bookdown::pdf_document2:
    template: templates/brief_template.tex
  bookdown::html_document2: default
documentclass: book
---

@@

```{r cohortNumbers, message=FALSE,echo = FALSE}
```

@@

As part of a broader investigation into the relationship between lipid levels and dementia risk, I applied for access to `r n_applied$total` unique data sources containing information on these variables. However, only a small proportion of these data sources (`r fmt_applied(n_accessed$total)`) provided me with access.

The resulting analysis did not suggest an effect of any lipid fraction (total cholesterol, LDL-c, HDL-c, or triglycerides) on any dementia outcome. In addition, patients characteristics such as age and sex did not have any impact on the relationship of lipid fractions and dementia outcomes.

Finally, the reasons for the low response rate are explored and future studies to investigate how data access rates could be improved are suggested.

@@

Individual patient data meta-analyses are considered to be the gold standard form of evidence synthesis, allowing for the application of a common selection model and analytical approach across all identified cohorts. They are particularly useful for investigating the impact of patient-level characteristics, something that is not possible with aggregate data unless the results have been explicitly stratified by the characteristic of interest.

While the systematic review presented in Chapter \@ref(sys-rev-heading) did not provide strong evidence for an effect of age at lipid measurement or patient sex on the relationship between lipids and dementia, the number of included studies in the lipid fraction meta-analyses were very small, due to both the relatively small number of studies reporting on this exposure and the poor reporting of smmary statistics of patient characteristics.

As such, the aims of this analysis were two fold: firstly, to perform a individual patient data meta-analysis across identified cohorts to examine the impact of patient age-at-measurement and sex on the effect of lipids on dementia; and secondly, to expand the evidence base for the effect of lipids on dementia outcomes by incorporating previously unanalysed cohorts contained in the Dementia Platform UK, a large consortia of dementia cohorts.

@@

Eligible data sources were defined as those that contained lipid measurements and lipids reported/available as a continuous measure, Data sources which were cross-sectional, either by design or due to the available data (e.g. a study conducted multiple waves, but onyl data from a single wave could be accessed) were excluded. Similarly, due to the limited time-frame, studies making use of population-level electronic health records, which often require an extensive project proposals in order to gain access to the data, were ineligible due to the time and cost involved in applying. The one exception to this was data from the CPRD, which I already had access to as part of the analysis reported in Chapter \ref(cprd-analysis-heading).

@@

Potentially eligible data sources were identified via two mechanisms with a distinct focus. For each approach, the number of cohorts responding to the request, and where applciable, the reasons given for a lack of data sharing were recorded.

#### Systematic review

The first approach focused on previously analysed data sources. The systematic review detailed in Chapter \@ref(sys-rev-heading) identified several observational prospective cohort studies examining the effect of blood lipid levels on dementia outcomes. The data sources used in each of these cohort analyses were screened against the criteria listed in the previous section, and eligible cohorts were approached for data access. In the first instance, the first/corresponding author of the publication was emailed (see Appendix \@ref(ipd-email-collab) for a copy of the text and documentation sent to potentially collaborators) in Autumn 2020. If this approach did not elicit a response within two months, the last author was contacted, on the basis that the first/corresponding author may have been a more junior member of the research group who has moved to a different institution.

#### Dementia Platform UK

The second approach focused on incorporating previously unanalysed data, thus providing additional evidence on the relationship between statins and dementia risk.

This was achieved through the Dementia Platform UK (DPUK), a collaborative grouping of existing dementia cohorts established by the MRC which works with data owners to make their data readily accessible for secondary analysis.[@bauermeister2020] It provides access to 42 cohorts with over 3 million participants, and makes use of a central streamlined application processs for all cohorts with the intent of making it easier to access data from existing data sources.

Cohorts included in the DPUK were assessed against the eligibility criteria, and in Autumn 2020, an application for access to a subset of 16 cohorts was made via the common-access procedure.

@@

Due to the inherent conflict of interest in assessing my own analysis, the result were also assessed by two independent reviewers.

@@

@@

### Variable definition

#### Exposure

To aid interpretability of the outcome, all analyses were standardised to refer to a 1-standard deviation increase in the lipid fraction under investigation.

#### Outcomes

Similar to the analyses presented in othe chapters, th endpoints in this analysis was incidience of dementia and related sub-types (AD, vascular dementia).

#### Covariates

Relevant covariates were adjusted for in all models. these included age, sex,education and comorbities. For information on how these were defined across cohorts, see Appendix \@ref(hold).

### Statistical analysis

#### Descriptive statistics

All IPD obtained was examined and recoded as necessary to ensure they comply with the coding guideline for the IPD database. As recommended , summary statistics for each data source were calculated and compared to publicly available statistics to ensure no errors were introduced in the data cleaning process.[@levis2021] Any discrepancies were investigated.

#### Data cleaning and harmonisation

Across all cohorts, data cleaning was performed in a similar manner, using commonly named variables, so that a single model could be applied using functional programming. The one exception to this is the CPRD data, which was held in a different system to the rest. However, an identical model was applied within this cohort. The advantage of this approach is that it reduces the likelihood of errors in model mis-specification if needing to change variables names from cohort to cohort.

#### Individual patient data analysis

<!-- Need detailed paragraph here describing the different between two stage versus one-stage approach -->

In this analysis, an two-stage individual patient data model was used out of necessity, with the one-stage model being precluded by the included datasets being in protected data silos. It also allows for each studys' data to be analysed according to a common analysis plan, which is particularly important in this topic, due to the differing definitions between studies of the serum concentrations above or below which lipids are deemed to be at abnormal levels.

Under a two-stage approach, estimates for the effect of each lipid fraction on available dementia outcomes in each study were first calculated for each data source. Effect estimates for the associations between lipid fractions and incident dementia, adjusted for identical covariates, were obtained by applying a common model to each data source. More specifically, hazard ratios were used to quantify the effect of a 1-SD lower exposure to a lipid fraction on dementia outcomes. A discrete proportional hazards model was employed to account for the interval censoring introduced by design of the longitudinal cohort studies.[@wang2017] An overall effect estimate was then produced by combining data-source-specific estimtes in a random-effects meta-analysis (see Section \@ref(hold) in Chapter \@ref(hold) for a broader discussion of meta-analysis methods). Following best practice, clustering within studies was accounted for.[@abo-zaid2013]

@@

#### Investigating effect of patient-level covariates

@@

#### Combining published aggregate data and individual participant data

Where a data source did not provided their individual participant data, but stratified by a patient characteristic of interest in the published data, the aggregate data and individual data were combined.

<!-- TODO How does this work? -->

### Comparison with published findings

Where an included data source had previously been analysed and results for the association between lipid levels and dementia outcomes reported, these were compared to the results of this analysis. If discrepancies were identified, these were investigated.

@@

### Secondary analyses

#### Amount of information added

The proportion of new information added to the evidence base via incorporation of the unanalysed datasets was quantified, by addding the additional results to the lipid-dementia meta-analyses presented in Section \@ref(hold), and extracting the combined weights assigned to the new data in a fixed-effect model.

<!-- #### Contact information  -->

#### Data availability statements

@@

Of the `r n_applied$total` studies to which I applied for data access, only three (`r paste0(comma((n_accessed$total/n_applied$total)*100),"%")`) provided data and were included in the analysis. Figure \@ref(fig:cohortFlowchart) details whether the cohorts eventually included in the review were identified by the systematic review or via DPUK portal. In addition, the reasons for cohorts not being included in the analysis are presented, stratified by application approach.

In summary, the requests for data from cohorts identified by the systematic review was characterised by a very low response rate. For the minority who did respond, common reasons given by corresponding authors for not sharing the data included that they no longer worked with the same group and did not know if the data were available or how to obtain it, had access to the data, or that they were currently or intended to perform a similar analysis as the one proposed.

With respect to the application to DPUK cohorts, where a dedicated project manager liaises with data owners on the applicants' behalf, the overall response rate was higher resulting in eventual access to three cohorts (one of which was also identified by the systematic review). However, even using this streamlined approach, a response was obtained for only half of the approach cohorts.

@@

As highlighted in Figure \@ref(fig:cohortFlowchart), there was little overlap between cohorts identified by the systematic review and those contained in the DPUK (`r fmt_applied(n_applied$DPUK_sysrev)`), indicating that the DPUK is a useful source of unanalysed data with respect to this question.

@@

The four data sources used in this analysis are summarised in Table \@ref(tab:cohortOverview-table) and are described in detail in the following sections. Of note, all included data sources were based in the United Kingdom. This is likely due to the majority of included datasets coming from the Dementia Platform UK route (Figure  \@ref(fig:cohortFlowchart)), which as implied by the name, has a geographical focus on cohorts performed in the United Kingdom.[@bauermeister2020]

@@

The main reason for exclusion within this records was use of an electronic health record database. Examples included the Taiwan National Health Insurance database and studies run in the US making use of Medicare data.

@@

Several additional cohorts are present in the DPUK collection but were not formally approached with a request for data access. This was primarily due to the information obtained via the online data dictionary system or via initial contact with the cohort owners indicating that the study was relatively new (and so was yet to collect more than 1 wave of data) or that the exposure variables of interest were not recorded.

<!-- TODO LBC 1936 -->

@@

As highlighted in Figure \@ref(fig:cohortFlowchart), several cohorts from the DPUK responding positively but on inspection of data provided these cohorts were excluded for a range of reason. The exclusion reason for each cohort for which data were provided is illustrated in Table \@ref(tab:dataExcluded-table).

@@

<!----------------------------------------------------------------------------->
(ref:dataExcluded-caption) Exclusion reasons for cohorts providing data

(ref:dataExcluded-scaption) Exclusion reasons for cohorts providing data

@@

BRACE is a memory-complaint-based cohort, suggesting that it is unlikely they were dementia-free at baseline). <!-- TODO CITATION NEEDED --> Similarly, the Memento study was excluded as it's criteria for entry required participants to be outpatients from a memory clinic. <!-- TODO CITATION NEEDED --> The Generation Scotland provided only cross-sectional data only for the variables of interest, as did the the NICOLA study. <!-- TODO CITATION NEEDED x2 --> Finally, the ELSA study was excluded based on the definition of the exposure variable. This study provided a "high cholesterol" a binary variable rather than the raw continuous lipid measurement data. <!-- TODO CITATION NEEDED -->

@@

This was particularly common among the cohorts identified via the systematic review and included studies with  a dedicated access process or data access committee. A cohort of particular interest, having been used in multiple studies identified by the systematic review, is the Three City Study in France. Despite multiple attempts to contact the team, there was no response received. <!-- TODO CITATION NEEDED --> This study was approach at multiple points over the curse of this analysis, and no answer was received.

@@

### Analytical results

There was no evidence for an effect of any lipid fraction on any dementia outcome (Figure \@ref(fig:tcDementiaMain)). Additionally, meta-analysis of the interaction coefficients for age-at-lipid-measurement and participant sex did not reveal an evidence for an varying effect across these variables.

@@

@@

@@

<!-- ### Effect by age -->

<!-- ### Heterogeneity -->

@@

### Summary of findings

This analysis requested data from `r n_applied$total` data sources, but only obtained data from `r fmt_applied(n_accessed$total)`, all of which were based in the United Kingdom. No evidence for an effect of lipids on the risk of dementia or related outcomes was identified. Similalarly, there was no evidence for an effect of age-at-measurement or sex on the lipid-dementia relationship.

Additionally analyses revealed that authors of cohorts indentified by the systematic review rarely stated that the data underpinning their analyses would be freely available, while a review of corresponding authors of these publications demonstrated that for many, the institutional email address provided was out of date.

<!-- Useful citation for discussion[@levis2021] -->

@@

<!-- TODO Plot showing when datasets requested were published and comparing this with the -->
A key limitation of this analysis is the very low response rate to requests for data access. This finding is not unexpected however, given that a review of IPD studies published between 1987 and 2015 found that fewer than half managed to obtained data from greater than 80% of studies, <!-- TODO CITE and are these RCTs or all studies? --> and that in many cases, the exact percentage of studies for which data were obtained was not accurately reported.

There are likely several likely reasons for this low response rate. In general terms, there are a several well-documented reasons why data are not made readily available, including concerns regarding participant privacy, concerns around scoping or "parasitic" behaviour, and a lack of trust between primary and secondary researchers. <!-- TODO CITATION NEEDED -->

More specific to this analysis, individual participant data meta-analysis including studies other than randomised controlled trials have less success in obtaining individual participant data from studies.[@nevitt2017a] Additionally, while no evidence is available on whether the characteristics of the researcher requesting data access influences the response rate and eventual decision, there is the possibility that my position as a PhD student meant I was less likely to elicit the response rate than a well-known professor might. The timing of the requests for data access, coinciding with a global pandemic, may also have affected the response rate as researchers prioritised other work.

_Many authors claim that their data can be made “available on request”, despite previous work establishing that these statements are demonstrably untrue in the majority of cases - that when data is requested, it is not actually made available. [@krawczyk2012; @miyakawa2020; @naudet2018] Additionally, previous work found that the availability of data “available on request” declines with article age, indicating that this approach is not a valid long term option for data sharing.[@vines2014] Several of the included studies are greater than 10 years old, and so obtaining data from these cohorts is less likely compared with more recently published studies._ <!-- TODO taken from DAS paper, so will need to change language -->

One of the reasons for this may be that the email addresses reported on publications are more likely to be out of date for older publications. Qualitatively, investigation of a subset of cohorts revealed that several corresponding/first authors were no longer at the same institution as when the study was reported, and as a result, were unlikely to have access to the institutional email address listed on the study publication. <!-- All of these are arguments towards better reporting of data availability, ownership - also talk about the ORCID and advatnages of permanent identifier-->

 These challenges are in theory what the DPUK was built to address. However, even with the help of the streamlined application process afforded by the DPUK, accessing sufficient data was a challenge. The response rate among DPUK cohorts a year after application was just 50%. In addition, some cohorts responded that that the proposed study question was already under investigation by another group, and that they would not share the data on this basis. In light of this, a centralised database of ongoing analysis being performed in DPUK would be of enormous help. Finally, the DPUK should more clearly distinguish between those cohorts that are "DPUK native" versus externally hosted, as the response time for externally hosted cohorts is likely to be much longer.

<!-- TODO CITATION NEEDED: paper on trust as a big issue + paper on systematic review of barriers to data sharing -->

Whether or not to press ahead with an IPD analysis in the absence of all (or even most) data is a personal decision, and some previous analyses have highlighted where they decided not to pursue an IPD study.[@jaspers2014]

@@

For the purposes of this thesis, conducting the IPD analysis was useful as it afforded me the opportunity to gain experience in applying the methodology. In addition, it allowed me to analysed two addition previously unanalysed cohort studies which were then utilised in the triangulation study detailed in Chapter \@ref(tri-heading).

@@

<!-- #### Opportunistic inclusion of CPRD data -->

@@

While this analysis did not manage to systematically obtain and analyse a large proportion of identified data, it did enable an analysis of two previously unanalysed datasets - the CaPS study and the EPIC Norfolk study - providing additional data that is incorporated into the triangulation exercise reported in Chapter \@ref(tri-heading).

@@

<!-- ### Study within a review -->

<!-- In future, it could have been worth running a "study within a review", such as that used to examine the best method [@godolphin2019a] to identify and access study data. -->

<!-- Options could include an investigation of the impact of applicant chacateristics  -->

<!-- Previous SWARs have been registered as protocols, but have not yet reported! -->

@@

In hindsight, attempting to undertake a large-scale IPD meta-analysis as part of a larger PhD project may have been overly ambitious. Data harmonization between cohorts in an IPD analysis is an often under-appreciated challenge,[@levis2021] and in line with this, data cleaning for this analysis took substantially longer than expected.

<!-- While the cohort response rate was substantially lower that expected, given the time and resource intensity of data cleaning and harmonisation for just 4 cohorts,  -->

@@

## Conclusion

This analysis provided very weak evidence for an effect of lipid fractions on dementia outcomes. However, it does provide an analysis two previous unanalysed data sources that are subsequently used in Chapter \@ref(tri-heading).

### 2021-10-01 06-Chp-Triangulation.Rmd
Back to working version General clean-up Proper counting of words included in tables
End of day:

---
#########################################
# options for knitting a single chapter #
#########################################
output:
  bookdown::pdf_document2:
    template: templates/brief_template.tex
  bookdown::word_document2: default
  bookdown::html_document2: default
documentclass: book
---

@@

>>>>

---
bibliography: bibliography/references.bib
csl: bibliography/nature.csl
output:
  bookdown::pdf_document2:
    template: templates/brief_template.tex
  bookdown::word_document2: 
      toc: false
      toc_depth: 3
      reference_docx: templates/word-styles-reference-01.docx
      number_sections: false 
  bookdown::html_document2: default
documentclass: book
---

@@

While the triangulation framework provides for inclusion of multiple distinct types of evidence, based on the scoping search, it is likely that the three main sources will be RCTs of statins, multivariable regression of prospective cohorts, and the use of genetic instruments in observational data (Mendelian Randomisation studies). One important difference between these approaches that will need to be addressed in the triangulation
is that RCTs relate to short-term modification of lipid levels, whereas Mendelian randomisation studies typically refer to lifelong exposures; and traditional observational epidemiology approaches may refer to short or longterm exposure (although are often not explicit about this). The triangulation will take these into account, for example by superimposing the studies on a hypothetical model for exposure-outcome relationship across the life course.

### 2021-10-01 front-and-back-matter/99-appendices.Rmd
End of day:

## Chapter \@ref(ipd-heading) {#appendix-ipd-analysis}
<!----------------------------------------------------------------------------->
### Section 1

>>>>

## Chapter \@ref(ipd-heading) {#appendix-ipd-analysis}
<!----------------------------------------------------------------------------->
### Email and documents sent to potental collaborators {#ipd-email-collab}
